\chapter{ %Memory Observable Injectable  Startable Trace 
MOIST simulations and semantics of CompCert}\label{ch:compcert}

The simple code in \autoref{example:remember} communicates with its environment in two main ways: (1) it takes an address as input and (2) reads from and writes to this location to increment the value stored there.
\begin{figure}\begin{lstlisting}[language=C, numbers=left]
int *buff;
void remember(int *p){
    buff = p;
}
void incr(void){
    ++(* buff);
}
\end{lstlisting}
\caption[Code example: \Ccode{remember} and \Ccode{incr}]{The function \Ccode{remember} records the address of some buffer, and \Ccode{incr} increments it by one. }
\label{example:remember}
\end{figure}
We will see how the specification of CompCert prevents us from reasoning about such programs, either as compilation units or as external functions, and we will show how to extend the specification of a compiler to lift these limitations.

First, CompCert can't give any guarantees about compiling the code in \autoref{example:remember} because it is not a complete program. It is reasonable to expect that the function \Ccode{remember} runs safely, given some assumptions (e.g. \Ccode{*p} is a valid address in memory). Unfortunately, CompCert's semantics assumes that a program starts executing with a call to  \Ccode{main()} with no arguments. CompCert characterizes the initial state by a predicate \Coqcode{initial_state: state -> Prop} that takes no additional arguments. You can see an instantiation of the predicate for Clight in \autoref{code:initial_state}. So, even though CompCert correctly compiles the code, its specification gives no guarantees of any execution other than the one that starts by calling main with no arguments.
\begin{figure}
\begin{lstlisting}
Inductive initial_state (p: program): state -> Prop :=
  | initial_state_intro: forall b f m0,
      let ge := Genv.globalenv p in
      Genv.init_mem p = Some m0 ->
      Genv.find_symbol ge p.(prog_main) = Some b ->
      Genv.find_funct_ptr ge b = Some f ->
      type_of_fundef f = Tfunction Tnil type_int32s cc_default ->
      initial_state p (Callstate f nil Kstop m0).
\end{lstlisting}
\caption[\Ccode{initial_state} predicate for C]{The \Ccode{initial_state} in C and Clight describes a call to \Ccode{main}. It also enforces that it takes no arguments (\Coqcode{Tnil}) and returns an integer (\Coqcode{ type_int32s}).}\label{code:initial_state}
\end{figure}

Second, imagine that the example in \autoref{example:remember} describes not the program being compiled but the semantics of two system calls \textsf{remember} and \textsf{incr}.  Suppose CompCert compiles some program that calls \Ccode{incr()}; then the compiler's specification gives no guarantee about the behavior of the compiled code. Indeed, CompCert's semantics allows calls to external functions that are assumed to be correct but, unfortunately, that specification of correctness is too strict; it  assumes that the function's behavior is fully determined by (1) the state of memory, (2) the function arguments and (3) the events produced by the function.\footnote{Leroy \cite{Leroy-Compcert-CACM} claims that ``inputs given to the programs are uniquely determined by their previous outputs", but this is not exactly correct. A more accurate representation of CompCert?s specification would be to say ``inputs given to the programs are uniquely determined by their most recent outputs". Indeed, the semantics of external calls \Coqcode{extcall_sem : Type :=  env -> list val -> mem -> trace -> val -> mem -> Prop} are determined by the environment, the arguments to the call, the current memory and produce a trace, a return value and a return memory. As we will see in \autoref{sec:compcert-sim}, it would be much stronger to determine inputs based on all historic outputs.} The behavior of \Ccode{incr} also depends on the value in \Ccode{buff} (which for system calls will not be in the program's accessible memory), so it is not correct, according to CompCert's specification. Certainly, \Ccode{incr} could expose the pointer stored in \Ccode{buff} as part of its trace but CompCert events, shown in \autoref{code:oldevents}, can only contain integers, floats or pointers to global variables.
 
\begin{figure}\centering
\begin{multicols}{2}
\begin{lstlisting}[style=CoqTheorem-list]
Inductive event: Type :=
  | Event_syscall: 
  	string -> list eventval -> 
  	eventval -> event
  | Event_vload: 
  	memory_chunk -> ident -> 
  	ptrofs -> eventval -> event
  | Event_vstore: 
  	memory_chunk -> ident -> 
  	ptrofs -> eventval -> event
  | Event_annot: 
  	string -> list eventval -> event
  \end{lstlisting}
  \columnbreak
  \begin{lstlisting}
Inductive eventval: Type :=
  | EVint: int -> eventval
  | EVlong: int64 -> eventval
  | EVfloat: float -> eventval
  | EVsingle: float32 -> eventval
  | EVptr_global: ident -> ptrofs -> eventval.
  \end{lstlisting}
\end{multicols}
\caption[Events in CompCert]{The events in CompCert}\label{code:oldevents}
\end{figure}

Moreover, in the CompCert semantics, the entire behavior of each external function call is bundled into one big step. Looking at an execution, internal steps and external function calls are uniform. This consistency is very useful when reasoning about the compilation of the program, where we want to abstract external calls. Nevertheless, when reasoning about a program in a context, it is more useful to replace the big step external calls with their small step semantics. Regrettably, the specification of CompCert does not even guarantee that the source and target programs call the same external functions. In theory, CompCert could replace an external function call with internal steps as long as they had the same (possibly empty) trace. In practice, obviously, CompCert does not do that, this guarantee is not exposed in its specification.

Finally, the correctness of CompCert is stated as a semantic preservation theorem, where the traces are the preserved behavior and the proof uses forward simulations\footnote{Forward simulation and determinism of the target language implies bisimulation and thus preservation of behavior.}. At least two other works (\cite{compcomp}, \cite{jiang14:pldi}) have proposed alternative CompCert specifications that make the simulations an exposed feature of the compiler's specification. In these papers, the authors view CompCert correctness modularly as a thread-local or module-local simulation and recover a simulation of the global program later. Moreover, from the exposed simulations, they can recover the relation between memories in source and target, another very useful feature in compositional compilers, which seems to be a key feature in compositionally. Following this line of work, we propose to expose the simulation as the specification of the compiler, deriving semantic preservation as a corollary. 

Finally, in each CompCert intermediate language, every state of the small-step semantics contains exactly one memory.  To reason parametrically about how a thread of execution evolves its memory, we need a uniform way to access the memory component of the small-step state.
 
We move, then, to lift these limitations according to the following richer notion of specification:
\begin{definition}[MOIST simulations]
We say that a compiler's specification uses Memory, Observable, Injectable and Startable Trace (MOIST) simulations if they satisfy the following:
\begin{itemize}
\item Memory: All intermediate languages have a unified memory model \Coqcode{mem}, and each language $L_1$ has a function \Coqcode{get_mem: state L\_1 -> mem}, that exposes the memory of a state. The simulation describes the relation between memories before and after compilation.  
\item Observable: Similarly, all intermediate languages are outfitted with a function \Coqcode{at_external} that identifies states about to make an external function call. For every language $L_1$, \Coqcode{at_external: state L_1 -> option (f_ext, args)} returns the external function being called and its arguments. The simulation preserves external calls. 
\item Injectable: The execution trace supports events that can describe locations in memory (i.e., pointers). Compilation may rearrange memory, which CompCert describes as an \emph{injection}, so the trace will be preserved up to these injections. The simulation shows that the injection relating traces in source and target executions is the same injection that relates their memory. We call these new events \emph{memory events}.
\item Startable: The execution of a program can start in any of its public functions, including \Ccode{main}, taking arguments.  

%\item The correctness of the compiler is stated as a forward simulation and the semantic preservation derive as a corollary. The forward simulations preserve external function calls and exposes the relation between source and target memories, throughout the execution. We call this an \emph{exposed simulation}. 
\end{itemize}
\end{definition}
It is worth noting that, even though we require a unified memory model, in practice, a language can use a different memory model (or none at all) as long as it can construct a memory from its state with \Coqcode{get_mem}. In practice all CompCert languages use the same memory model, described in \autoref{sec:compcert-mem}, which we will refer to as \Coqcode{mem} from now on.
Nevertheless, a future language could use \emph{juicy memory} as in \cite{appel14:plcc} or abstract state in \cite{Ronghui15abstractionlayers} since a \Coqcode{mem} can be derived from them. 

\def\percchange{3.6\%}
\begin{table}
\centering
\begin{tabular}{|c|c|c|} \hline  
		& Percent change & Number of lines changed \\
\hline Arguments in \Ccode{main} & 1.6\% & 3660 \\ %Calculated as diff, to when entry_points branch started: gif diff --stat ff3f7c830c7676f7847ca9391d2786da3236c9f6
\hline Injectable Traces & 0.5\% & 1141 \\  %use [ git diff --stat f047fcb7852ff58c0c62f10d41f91f3f88552780] and look changes in [Events.v + EventsAux.v]
\hline MOIST Semantics & 0.2\% & 466 \\  % [387 = 43 * 9 + 79] calculated as number of lines to define a new semantics * number of languages. + new deifnitions of semantics.
\hline MOIST Simulations & 1.3\% & 2911 \\ % The rest: = 8178 - 466 - 1141 - 3660 - 2911
\hline \textbf{Total} & \percchange & 8178 \\ %total computed with  git diff --stat f047fcb7852ff58c0c62f10d41f91f3f88552780. Removing the lines from Driver/compiler.v
\hline \end{tabular}
\caption[Measures of changes to CompCert.]{Measures of changes to CompCert: changes are calculated from the number of lines added as given by running \Ccode{git diff} between our code and the commit of CompCert we branched off from. For each feature, an estimated number of lines is provided. }\label{tab:prcntchange}
\end{table}

In the rest of the chapter, we describe how we develop MOIST specifications for CompCert. We first describe how to generalize \Coqcode{initial_state} to make the simulations Startable. Second we describe how to add memory events to CompCert. Then we show how to extend the semantics for every language in CompCert to include \Ccode{get_mem} and \Coqcode{at_external} functions and, finally, we show how to put everything together in MOIST simulations for CompCert, which proves the following theorem:

\begin{theorem}[CompCert MOIST specification]\label{thm:moistsim}
Given a source program P and an x86 assembly language program Q obtained by CompCert compilation of P, the CPMs of the two programs are in a MOIST simulation relation.
\end{theorem}
  

The changes described here represent only a \percchange\ %TODO
change to CompCert, as measured by running \Coqcode{git diff} in CompCert before and after our changes. The amount changed for every feature proposed is described in \autoref{tab:prcntchange}.\footnote{All these measurements are made with respect to a commit in CompCert from May 21 (\Ccode{f047fcb7852ff58c0c62f10d41f91f3f88552780})}
Approximately 10\% of all new additions are specifications and the rest are proofs. 

\input{ch-compcert/compcert-memory}
\input{ch-compcert/compcert-main_args}
\input{ch-compcert/compcert-events}
\input{ch-compcert/compcert-semantics}
\input{ch-compcert/compcert-simulations}
